---
title: "Assignment 7 - Answers"
author: "Charles Lang + Yiran Du"
date: "11/30/2016"
output: html_document
---

In the following assignment you will be looking at data from an one level of an online geography tutoring system used by 5th grade students. The game involves a pre-test of geography knowledge (pre.test), a series of assignments for which you have the average score (av.assignment.score),  the number of messages sent by each student to other students about the assignments (messages), the number of forum posts students posted asking questions about the assignment (forum.posts), a post test at the end of the level (post.test) and whether or not the system allowed the students to go on to the next level (level.up).  

#Upload data
```{r}
D1 <- read.csv("~/Desktop/github/assignment7-YiranDuMirabelle11/online.data.csv")
```

#Visualization 
```{r}
#Start by creating histograms of the distributions for all variables (#HINT: look up "facet" in the ggplot documentation)
library("ggplot2")
library("reshape2")
library(corrplot)
D1 <- D1[-1]
D1$level.up <- ifelse(D1$level.up == "yes",1,0)
head(melt(D1))
ggplot(data = melt(D1), mapping = aes(x=value)) + geom_histogram(bins = 10) + facet_wrap(~variable, scales = 'free_x')
#Then visualize the relationships between variables
pairs(D1)
D1COR <- cor(D1)
corrplot(D1COR, method="circle", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")
```
#### Try to capture an intution about the data and the relationships:
We found in the visualization that all pairs in our variables are positively correlated. Some of the variables are strongly correlated, e.g. the number of messages and their post-test score. Therefore, it seems that the relationship among this dataset is strong.



#Classification tree
```{r}
#Create a classification tree that predicts whether a student "levels up" in the online course using three variables of your choice (As we did last time, set all controls to their minimums)
library("rpart")

## Since in the graph above, we find that the first three variables strongly correlated with level up is post-test score, messages and average assignment score, we will adopt these three in our first model
rp <- rpart(level.up ~ post.test.score + av.assignment.score + messages, method="class",control = rpart.control(minsplit = 1,minbucket= 1, cp=0), data=D1)
printcp(rp)

#Plot and generate a CP table for your tree 
post(rp, file="tree1.ps", title = "Tentative Classification of Level Up over post-test-score, average assignment grade and messages")

#Generate a probability value that represents the probability that a student levels up based your classification tree 

D1$pred <- predict(rp, type = "prob")[,2]#Last class we used type = "class" which predicted the classification for us, this time we are using type = "prob" to see the probability that our classififcation is based on.

#Now you can generate the ROC curve for your model. You will need to install the package ROCR to do this.

library(ROCR)

#Plot the curve
pred.detail <- prediction(D1$pred, D1$level.up) 
plot(performance(pred.detail, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(pred.detail,"auc"), "y.values"))#Unlist liberates the AUC value from the "performance" object created by ROCR



#Now repeat this process, but using the variables you did not use for the previous model and compare the plots & results of your two models. Which one do you think was the better model? Why?

# The model with the rest two variables plus messages:
rp2 <- rpart(level.up ~ pre.test.score +forum.posts + messages, method="class",control = rpart.control(minsplit = 1,minbucket= 1, cp=0), data=D1)
printcp(rp2)

#Plot and generate a CP table for your tree 
post(rp2, file="tree2.ps", title = "Tentative Classification of Level Up over pre-test-score, forum posts and messages")

#Generate a probability value that represents the probability that a student levels up based your classification tree 

D1$pred2 <- predict(rp2, type = "prob")[,2]

#Now generate the ROC curve for the model.

#Plot the curve
pred2.detail <- prediction(D1$pred2, D1$level.up) 
plot(performance(pred2.detail, "tpr", "fpr"))
abline(0, 1, lty = 2)

#Calculate the Area Under the Curve
unlist(slot(performance(pred2.detail,"auc"), "y.values"))

# The first one is better since it has a high AUC, which approaches 1.
```
#Thresholds
```{r}
#Look at the ROC plot for your first model. Based on this plot choose a probability threshold that balances capturing the most correct predictions against false positives. Then generate a new variable in your data set that classifies each student according to your chosen threshold.
diagn <- function(threshold) {
  
D1$threshold.pred1 <- D1$pred> threshold

#Now generate three diagnostics:

D1$accuracy.model1 <- sum(D1$threshold.pred1)/nrow(D1)

D1$false.positive <- D1$threshold.pred1 == TRUE & D1$level.up==0
D1$truepostive <- D1$threshold.pred1 == TRUE & D1$level.up ==1
D1$precision.model1 <- sum(D1$truepostive) / (sum(D1$truepostive) + sum(D1$false.positive))

D1$flse.negative <- D1$threshold.pred1 == FALSE & D1$level.up==0
D1$recall.model1 <- sum(D1$truepostive) / (sum(D1$truepostive) + sum(D1$false.negative))

#Finally, calculate Kappa for your model according to:

#First generate the table of comparisons
table1 <- table(D1$level.up, D1$threshold.pred1)

#Convert to matrix
matrix1 <- as.matrix(table1)

#Calculate kappa
predict.kappa <- kappa(matrix1, exact = TRUE)/kappa(matrix1)

return(predict.kappa)
}

diagn(0.8)
#Now choose a different threshold value and repeat these diagnostics. What conclusions can you draw about your two thresholds?
# We set the threshold to 0.5
diagn(0.5)

# The kappa does not change because in our model(We set all the controls to its minimum and AUC = 1), the model is perfect to our dataset at hand, though with the problem of overfitting and the threshold will not change our kappa.  But if we loosen the control and re-calculate the kappa. We will find that kappa itself cannot determine whether a model is good or not since kappa will decrease with a higher threshold.

```


